\toolname: Property-Based Testing for MPC Security

### 2023-11-03 paper/main.tex
knockinng out TODOs

At face value, property-based testing doesn't generally extend to hyper-properties,
which can only be assessed by comparing multiple program traces.
In practice sometimes one can express a hyper-property as a property of multiple of duplicate programs,
but the sensitivity of the test quickly decays and the logistics of setting up the test quickly grow.
Furthermore, when the property to be tested is over distributions of traces,
the various used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.

@@

Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
use EasyCrypt to manually show the security of GMW.
Their definitions generalize to protocols with efficiently invertible functionalities.
% remove? Since existing general-purpose SMPC protocols can easily compute non-invertible functions, this is a serious limitation.
They model SMPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
If their system could be automated and paired with language constraints to ensure invertibility then it
could apply to many, but certainly not all, use cases.

@@

Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
give a powerful language system that can check if one protocol is a valid simulator of another,
and show that it works on existing MPC protocols like GMW.
The user of the system needs to write a second protocol
representing the composition of the ideal-world functionality and the explicit function $\mathtt{SIM}$,
and then build a high-level proof of the equivalence in the provided equational logic.

The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
give fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
[TODO: there are other/later works along these lines we should cite?]

Darais et al.~\cite{darais2019language} %
gives a type system that ensures uniform and independent randomness of key values throughout a program.
Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
but this would of course prevent any non-vacuous functionality from being implemented.

OWL~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
They can type-check more convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
there's no way to verify an implementation of MPC.

Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
also use a type system to ensure safe usage of cryptographic primitives,
but they specifically target homomorphic encryption schemes and the various measures by which such schemes may be correct.
Like OWL, their type system assumes a secure implementation of the encryption scheme.

@@

We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
A SPN is directed-graph representation of a multi-variate distribution;
they are compact, allow efficient inference, and can represent values symbolically as needed,
so testing for conditional independence is practical.
However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.

The DICE language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
but since the graph \textit{structure} of the BDD of a DICE program depends on the actual \textit{values} on which one is conditioning,
a dice-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.

Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
is a program logic for reasoning about probability distributions and PPLs.
It more expressive than the original probabilistic separation logic,
but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.

@@

Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
Specifically, if decision trees (D-Trees trained on random variables $A$ and $C$ do better at predicting the value of a random variable $B$,
then $A$ and $B$ are not independent conditioned on $C$.
The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
but in our adaptation we use fresh program traces for everything.

@@

We use the techniques from \cite{chalupka2018fast}
to directly test for conditional dependence without any static analysis of the subject protocol.

To actually test a protocol, one uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
One of the two models has access to the real-world views, and the only has access to the ideal-world views.
In principal, \textit{any} ML classifier is ok to use; the choice is partly an engineering question
and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
Generating the data for these tests requires running the protocol many times.
The process is enumerated in Algorithm 1.
Note that the data must be understood as vectors/matrices/tensors of bits;
aggregating values into larger structures such as bytes or integers would at best require careful modifications to Algorithm 1.

@@

The usual trade-offs of significance testing apply.
Setting $\alpha$ too low will cause programs who's insecure-ness is only marginally discernible to the D-Trees
to pass the test.
Setting $\alpha$ higher will compromise one's certainty that a failing program truly is insecure.
There are three reasons why this is not a serious problem with the test, each worth considering as its own subtlety.

First, if the test returns \texttt{INSECURE},
then there's nothing wrong with running the test again to see if you were just unlucky.
This assumes that you have the resources (\eg{} compute time) to do so, which you might not.
Since more powerful tests are better able to detect problems,
if your compute-budget is fixed you're better off spending it all on a single bigger test.

Second, we observe that the transition of an insecure program between "appears secure" and "appears insecure"
as the power of the test increases is sharp,
and as the size of the test continues to increase the $p$-values quickly drop below any reasonable $\alpha$.
This can be frustrating; if you're unsure of the security of a protocol
then there's always a chance a modestly-more-powerful test might be conclusive.
On the other hand, it means that at any reasonable fixed test-size,
most of the protocols it can detect insecure will be \textit{unambiguously} insecure.

The third such mitigating factor is also a problem on its own.
For a secure program the observed $p$-values
will typically \textit{not} be sampled from the uniform distribution on $\mathbb{R}_{(0,1]}$,
and $\alpha$ is actually an \textit{upper bound} on the likelyhood of falsely flagging a protocol as insecure.
Contrary to casual descriptions of significance testing, this is not a problem with the statistics;
it's a result of the inequality in the testing hypothesis.
[TODO: footnote: Significance testing of an \textit{equality} based null-hypothesis will yield uniformly distributed $p$-values
when the null-hypothesis is true.]
Because the ML models have no knowlege of the importance or semantics of the bits they're trained on,
the mostly-random message- and tape-data provided to the real-world models makes it harder for them
to learn from the input- and output-data they share with the ideal-world models.
In other words, the additional data on which the real-world models are trained is fundementally a hindrance to them,
and only becomes advantageous if the protocol is "insecure enough" to overcome this hidrance.
A protocol might be truely insecure and not have the relevant conditional independence,
but still appear secure because the advantage an adversary could gain at guessing the honest inputs
is smaller than the advantage the ideal-world ML models gain from their more focused data.
This issue could be overcome by setting \texttt{trainN} sufficiently high;
the ideal-world models' scores will saturate before the real-world models' do.

@@

Our python script for actually performing the test consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

TODO: link to github or other archive?

@@

>>>>

At face value, property-based testing doesn't extend well to hyper-properties,
which can only be assessed by comparing multiple program traces.
In practice sometimes one can express a hyper-property as a property of multiple duplicate programs,
but the sensitivity of the test quickly decays and the logistics of setting up the test quickly grow.
Furthermore, when the property to be tested is over distributions of traces,
the techniques used by good property-based testing frameworks to efficiently find interesting cases are not appropriate.

@@

Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
use EasyCrypt to manually show the security of GMW.
Their definitions generalize to protocols with efficiently invertible functionalities.
They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
If their system could be automated and paired with language constraints to ensure invertibility,
then it could apply to many, but certainly not all, use cases.

@@

Gancher et al.~\cite{gancher2023core} % https://dl.acm.org/doi/pdf/10.1145/3571223
give a powerful language system that can check if one protocol is a valid simulator of another,
and show that it works on existing MPC protocols like GMW.
The user of the system needs to write a second protocol
representing the composition of the ideal-world functionality and an explicit $\mathtt{SIM}$,
and then build a high-level proof of the equivalence in the provided equational logic.

The \textsc{Wysteria} language~\cite{rastogi2014wysteria} % https://www.ieee-security.org/TC/SP2014/papers/Wysteria_c_AProgrammingLanguageforGeneric,Mixed-ModeMultipartyComputations.pdf
gives fine-grained integration of MPC as a language primitive, to help a larger program use it efficiently.
Their system assumes (at least one) secure implementation of MPC on circuits.

$\lambda_\textbf{obliv}$~\cite{darais2019language} %
gives a type system that ensures uniform and independent randomness of key values throughout a program.
Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
but this would of course prevent any non-vacuous functionality from being implemented.

\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
They can type-check convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
there's no way to verify an implementation of MPC.

Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
also use a type system to ensure safe usage of cryptographic primitives,
but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
Like OWL, their type system assumes a secure implementation of the encryption scheme.

@@

We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
A SPN is a directed-graph representation of a multi-variate distribution.
They are compact, allow efficient inference, and can represent values symbolically as needed,
so testing for conditional independence is practical.
However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.

The \textsc{Dice} language~\cite{holtzen2020scaling} translates programs into binary decision diagrams (BDDs),
analogous to the way we tried to translate them into SPNs.
BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.

Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
is a program logic for reasoning about probability distributions and PPLs.
It more expressive than the original probabilistic separation logic~\cite{barthe2019probabilistic},
but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.

@@

Our work builds directly on that of Chalupka et al.~\cite{chalupka2018fast}
who show that batches of ML classifiers (in practice, decision trees, because they're light-weight) can be used to check conditional independence.
Specifically, if decision trees (D-Trees) trained on random variables $A$ and $C$
do better at predicting the value of a random variable $B$ than D-Trees trained only on $C$,
then $A$ and $B$ are not independent conditioned on $C$.
The test is statistical; many D-Trees are trained and any apparent advantage of the better-informed ones must be assessed for statistical significance.
In their work the data is assumed to be permuted sub-samples of some finite pool of samples from a population,
but in our adaptation we use fresh program traces for everything.

@@

We use the techniques from \cite{chalupka2018fast} to build \toolname,
which directly tests for the above conditional dependence without any static analysis of the subject protocol.

To actually test a protocol, \toolname uses two identical machine learning (ML) tools to predict the honest secrets based on the corrupt views.
One of the two models has access to the real-world views, and the only has access to the ideal-world views.
In principal, \textit{any} ML classifier is ok to use; the choice is partly an engineering question
and partly a judgment call about what kinds of systems will be best able to discern patterns in the views of the subject protocol.
Generating the data for these tests requires running the protocol many times.
The process is enumerated in Algorithm 1.
Note that the data must be understood as vectors/matrices/tensors of bits;
aggregating values into larger structures such as bytes or integers would at best require careful modifications to line 8 in Algorithm 1.

@@

The usual trade-offs of significance testing apply.
Setting $\alpha$ too low will cause programs who's insecure-ness is only marginally discernible to the D-Trees
to pass the test.
Setting $\alpha$ higher will compromise one's certainty that a failing program truly is insecure.
There are three reasons why this is not a serious problem with \toolname, each worth considering as its own subtlety.

First, if \toolname returns \texttt{INSECURE},
then there's nothing wrong with running it again to see if you were just unlucky.
This assumes that you have the resources (\eg{} compute time) to do so, which you might not.
Since more powerful tests are better able to detect problems,
if your compute-budget is fixed you're better off spending it all on a single bigger test.

Second, we observe that the transition of an insecure program between "appears secure" and "appears insecure"
as the power of the test increases is sharp,
and as the size of the test continues to increase the $p$-values quickly drop below any reasonable $\alpha$.
This can be frustrating; if you're unsure of the security of a protocol
then there's always a chance a modestly-more-powerful test might be conclusive.
On the other hand, it means that at any reasonable fixed test-size,
most of the protocols \toolname \textit{can} detect insecure will be \textit{unambiguously} insecure.

The third such mitigating factor is also a problem on its own.
For a secure program the observed $p$-values
will typically \textit{not} be sampled from the uniform distribution on $\mathbb{R}_{(0,1]}$,
and $\alpha$ is actually an \textit{upper bound} on the likelyhood of falsely flagging a protocol as insecure.
Contrary to casual descriptions of significance testing, this is not a problem with the statistics;
it's a result of the inequality in the testing hypothesis.\footnote{
    Significance testing of an \textit{equality} based null-hypothesis would yield uniformly distributed $p$-values when the null-hypothesis is true.
}
Because the ML models have no knowlege of the importance or semantics of the bits they're trained on,
the (mostly random) message- and tape-data provided to the real-world models makes it harder for them
to learn from the input- and output-data they share with the ideal-world models.
In other words, the additional data on which the real-world models are trained is fundementally a hindrance to them,
and only becomes advantageous if the protocol is "insecure enough" to overcome this hidrance.
A protocol might be truely insecure and not have the relevant conditional independence,
but still appear secure because the advantage an adversary could gain at guessing the honest inputs
is smaller than the advantage the ideal-world ML models gain from their more focused data.
This issue could be overcome by setting \texttt{trainN} sufficiently high;
the ideal-world models' scores will saturate before the real-world models' do.

@@

Our \toolname python script consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

See Section~\ref{sec:public-repo} for more details.

@@

\section*{Data Availability}\label{sec:public-repo}

TODO.

This section, and everything following, does not count within the body text page limit requirement (so this paper would satisfy a page limit of 1). For your Euro S\&P submission, we expect more useful content than what is in this paper, so you have 13 pages of body text.
